---
title: "TBBT (Tidy)Text analysis 2 - TBD"
author: "Tamas Koncz"
date: '2018-04-02'
slug: tbbt-genders
tags:
- R
- tidytext
categories: text analysis
---

```{r setup, message=FALSE, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(knitr)
library(tidytext)
library(dplyr)
library(data.table)
library(stringr)
library(ggplot2)
library(gridExtra)
```

My inspiration for this post was Julia Silge post's on pudding.cool, titled ["She giggles, he gallops"](https://pudding.cool/2017/08/screen-direction/).  
Julia examines 2,000 film scripts, to uncover certain "gender bias" - notably, she compared what are the most frequent words following "he" and "she" in the scripts.  

I'm going to apply the same idea to the subtitle data of The Big Bang Theory, but, before we dive in - this post is the second in my "TBBT (Tidy)Text analysis" mini-series, if interested, you can find the first one [here](https://tomiajo.github.io/blog/tbbt-imdb-score-regression/), in which I look at the relation between average IMDB scores of episodes and dominating character "mentions".   
  

**And now, let's get rolling!** 
  
Data-wise, our starting point is the same as in the first post:
```{r, echo = FALSE}
rm(list=ls())

tbbt_subtitles <- readRDS("C:/Users/tkonc/Documents/Data/Text Mining Final Project/all_srts_df.rds")

full_episode_text <- tbbt_subtitles %>% 
                       select(episode_number, season_number, Text) %>%
                       mutate(episode_id = paste("s", season_number, 
                                                 "e", episode_number, sep = ""))

episode_order <- full_episode_text %>%
                   select(season_number, episode_number, episode_id) %>%
                   distinct() %>%
                   arrange(season_number, episode_number) %>%
                   select(episode_id) %>%
                   as.vector()

##creates 1 row / episode format                       
full_episode_text <- full_episode_text %>%
                       mutate(episode_id = factor(episode_id, levels = episode_order$episode_id)) %>%
                       group_by(episode_id) %>% 
                       summarize(episode_full_text = paste0(Text, collapse = " "))

full_episode_text %>%
  mutate(episode_full_text = paste(substr(episode_full_text, 1, 125), "...", sep = "")) %>%
  head(3) %>%
  kable()
```

The approach I'm going to take is simple: look at two-word pairs, where the first world defines the gender, and the second one is the word we are going to analyze.  
  
First, we need to slice the text into bigrams:
```{r}
bigrams <- full_episode_text %>%
             unnest_tokens(output = bigram, 
                           input  = episode_full_text, 
                           token  = "ngrams", 
                           n      = 2)
```

```{r, echo = FALSE}
bigrams %>%
  head(4) %>%
  kable()
```

I'll cross reference this bigrams list with a gender "lexicon" I have defined earlier, which apart from the usual "he/she", also contains the names of the important characters from the show (only first 6 lines shown below - for full list, please refer to my [github](https://github.com/tomiaJO/TEXT_MINING_FINAL_PROJECT/blob/master/gender_word_lexicon.csv)):  
```{r, echo = FALSE}
#### import gender lexicon ####
gender_word_lexicon <- fread("C:/Users/tkonc/Documents/Data/Text Mining Final Project/gender_word_lexicon.csv")

gender_words_regex <- paste0(gender_word_lexicon$word, collapse = '|')
gender_words_regex <- paste("^(", gender_words_regex, ") ", sep = "")

gender_word_lexicon %>%
  head(6) %>%
  kable()
```

Let's use this lexicon as a filter (via simple regex), and create separte columns for the two words of the bigrams as well:
```{r}
gender_bigrams <- bigrams %>%
                    filter(str_detect(bigram, gender_words_regex)) %>%
                    mutate(gender_word = str_split(bigram, pattern = " ", n = 2, simplify = TRUE)[,1],
                           word        = str_split(bigram, pattern = " ", n = 2, simplify = TRUE)[,2]) %>%
                    inner_join(gender_word_lexicon, by = c("gender_word" = "word")) %>%
                    select(gender, word)
```

What I am going to do exactly with these word pairs?  
  
My original idea was to run a tf-idf analysis, to "cluster" the most typical words for each gender.  
However, when I first ran a trial, I noticed something interesting: many words in the "she" category were sensual ones ("feel", "kissed", "smiled"...), while for "he" most words were pretty general ("talk", "sit", "boy"...).  
  
This _(and the sad fact that most of the words appeared only once or twice, making it hard to do meaningful tf-idf)_ made me pivot somewhat from my starting idea - and I decided to group words by sentiments, and compare those for the two genders.  
  
For this, I'm using the NRC lexicon, which is built-in accessible in the tidytext R package.
```{r, fig.align='center', fig.width=5, fig.height=3}
sentiments_nrc <- get_sentiments(lexicon = "nrc")

sentiments_nrc %>%
  group_by(sentiment) %>%
  summarize(n = n()) %>%
  mutate(sentiment = reorder(sentiment, n)) %>%
  ggplot(aes(x=sentiment, y=n)) +
    geom_bar(stat = "identity", fill = "springgreen4") +
    labs(title = "Words in the NRC sentiment lexicon",
         x     = "Word count",
         y     = "Sentiment") +
    coord_flip() +
    theme_minimal() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```
A glimpse into the sentiment lexicon's categories... The lexicon does not contain all English words, but should be more than enough for a light analysis of subtitles.

```{r}
gender_bigrams_w_nrc <- gender_bigrams %>%
                          inner_join(sentiments_nrc, by = "word")
```

In a short step we match up the sentiments with the bigram collection (above), and the results are ready to be visualized[^1]:

```{r, echo = FALSE, fig.width=15, fig.height=7, fig.align='center'}
gender_bigrams_w_nrc %>%
  count(gender, sentiment, sort = T) %>%
  tidyr::spread(key = gender, value = n, fill = 0) %>%
  ungroup() %>%
  mutate(male_total     = sum(male),
         female_total   = sum(female)) %>%
  mutate(male_ratio     = male / male_total,
         female_ratio   = female / female_total) %>%
  mutate(ratio          = male_ratio / female_ratio) %>%
  mutate(relative_ratio = 1 - ratio) %>%
  mutate(sentiment      = reorder(sentiment, -relative_ratio)) %>%
  mutate(direction      = ifelse(relative_ratio >= 0, "he", "she")) %>%
  mutate(he_sentiment   = ifelse(direction == "he",  toupper(as.character(sentiment)), ""),
         she_sentiment  = ifelse(direction == "she", toupper(as.character(sentiment)), "")) %>%
  ggplot((aes(sentiment, relative_ratio, fill = direction))) + 
    geom_bar(stat = "identity") +
    geom_text(aes(label = he_sentiment),  hjust =  1.05, 
                                          fontface = "bold", color = "grey50") +
    geom_text(aes(label = she_sentiment), hjust = -0.05, 
                                          fontface = "bold", color = "grey50") +
    labs(title = "Relative sentiment frequencies for the two genders",
         x     = "",
         y    = "") +
    scale_y_continuous(labels = c("1.5x", "1.25x", "same", "1.25x", "1.5x"), 
                       limits = c(-0.5, 0.5)) +
    scale_fill_manual(labels = c('Used more by "he"', 'Used more by "she"'),
                        values = c("deepskyblue4" , "brown1")) +
    coord_flip() +
    theme_minimal() +
    theme(panel.grid.major.y = element_blank()) +
    theme(axis.title.y = element_blank(),
          axis.text.y  = element_blank(),
          axis.ticks.y = element_blank()) +
    theme(plot.title   = element_text(size = 22, face = "bold", hjust = 0.5)) +
    theme(legend.title    = element_blank(),
          legend.position = "bottom",
          legend.text     = element_text(size = 12))
```

There is a high likelihood that "he" will be angry in the context mentioned - while "she" is mostly in the positive.

Footnotes:  

[^1]: describe methodology
